Reducing iptables configuration complexity using chains

### 2014-03-28 iptables_newline2014/presentation.tex
add Benefits of using chains
add chaining chains

  \begin{frame}
    \frametitle{Summary}
    \begin{itemize}
      \item use ESTABLISHED state to reduce number of rules
      \item using chains makes your rules easier readable and maintainable
      \item chains can be reused for several rules
      \item chains can be chained together
      \item faster, because it only jumps to a chain when a rule matches
    \end{itemize}
  \end{frame}
  \begin{frame}
   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?\\
    Contact
    \begin{itemize}
      \item Blog : \href{http://ruleant.blogspot.com/}{http://ruleant.blogspot.com}
      \item Twitter : \href{https://twitter.com/dcadriaenssens}{@dcadriaenssens}
    \end{itemize}
  \end{frame}
\end{document}

>>>>

    \frametitle{Benefits}
      \item the list of IP addresses in the custom chain is reused for both ports, so they have to be defined only once.
      \item adding/changing/removing an IP address is much easier
      \item a better overview of the firewall rules.
    \frametitle{Realworld example : webmaster PCs}
    Let's do the same for webmasters AND admins having access to SMB.\\
    Create a chain with webmaster IP addresses :
      \small{iptables -N webmaster\_IP\\
      iptables -A webmaster\_IP -s 10.100.2.11 -j ACCEPT\\
      iptables -A webmaster\_IP -s 10.100.2.17 -j ACCEPT\\
      iptables -A webmaster\_IP -s 10.100.2.34 -j ACCEPT\\
      iptables -A webmaster\_IP -j DROP}
    \frametitle{Realworld example : access to SMB}
    Add the rules for SMB (tcp port 139 and 445) using the custom chains :
      \small{iptables -A INPUT -p tcp -m tcp --dport 445 -j admin\_IP\\
      iptables -A INPUT -p tcp -m tcp --dport 445 -j webmaster\_IP\\
      iptables -A INPUT -p tcp -m tcp --dport 139 -j admin\_IP\\
      iptables -A INPUT -p tcp -m tcp --dport 139 -j webmaster\_IP}
  \textbf{This can be optimized!}
    \frametitle{Chaining chains}
    Include the sysadmin chain in the webmaster chain:
      \small{iptables -N webmaster\_IP\\
      iptables -A webmaster\_IP -s 10.100.2.11 -j ACCEPT\\
      iptables -A webmaster\_IP -s 10.100.2.17 -j ACCEPT\\
      iptables -A webmaster\_IP -s 10.100.2.34 -j ACCEPT\\
      \sout{iptables -A webmaster\_IP -j DROP}\\
      iptables -A webmaster\_IP -j admin\_IP}
    Jump to the sysadmin chain at the end of the webmaster chain.\\
    \textbf{Warning :} check conditions of chain ending
    \frametitle{Realworld example : access to SMB}
    The rules for SMB (tcp port 139 and 445) can be reduced to :
      \small{\sout{iptables -A INPUT -p tcp -m tcp --dport 445 -j admin\_IP}\\
      iptables -A INPUT -p tcp -m tcp --dport 445 -j webmaster\_IP\\
      \sout{iptables -A INPUT -p tcp -m tcp --dport 139 -j admin\_IP}\\
      iptables -A INPUT -p tcp -m tcp --dport 139 -j webmaster\_IP}
  The admin chain is checked as well because the webmaster chain includes it.

  \begin{frame}
    \frametitle{Conclusion}
    \begin{itemize}
      \item use ESTABLISHED state to reduce number of rules
      \item using chains makes your rules easier readable and maintainable
      \item chains can be reused for several rules
      \item chains can be chained together
      \item faster, because it only jumps to a chain when a rule matches
    \end{itemize}
  \end{frame}
  \begin{frame}
   \frametitle{Questions}
    Thanks for your attention!\\
    Questions?\\
    Contact
    \begin{itemize}
      \item Blog : \href{http://ruleant.blogspot.com/}{http://ruleant.blogspot.com}
      \item Twitter : \href{https://twitter.com/dcadriaenssens}{@dcadriaenssens}
    \end{itemize}
  \end{frame}
\end{document}
